Parameter estimation

### 2023-06-27 SFE-article/Article_Model_Parameters.tex
27.06

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ is the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
The aim of this study was to investigate the supercritical extraction process of the caraway oil from caraway seeds from a partially filled extractor with a fixed bed that was operated under multiple constant operating conditions. The dataset collected from the experiments consists of four experiments conducted at different operating conditions: $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar, and $50^\circ C$/$300$ bar. To describe the fluid-solid extraction process with $CO_2$ as a solvent, a distributed-parameter model is used. To reduce the number of spatial dimensions, the concept of quasi-one-dimensional flow is applied. It is assumed that the flow is uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, saturation concentration, and an initial state estimate. These parameters are estimated from the Peng-Robinson equation of state and by applying the maximum likelihood estimation method on yield data under the normal error assumption.

@@

The quasi-one-dimensional compressible Navier-Stokes equations in Cartesian coordinates are given by Equations \ref{EQ: CompressibleEuler_1} to \ref{EQ: CompressibleEuler_3}.

@@

{\footnotesize
	\begin{equation} \label{EQ: Pressure_Velocity}
		\cfrac{d v}{d {\color{orange}\rho_f}} = - \cfrac{d v}{d{\color{blue}P}} \cfrac{d P}{d {\color{orange}\rho_f}} = -\cfrac{1}{\rho v} \cfrac{d{\color{blue}P}}{d{\color{orange}\rho_f}} = -\cfrac{v}{{\color{orange}\rho_f}}
	\end{equation}
}

>>>>

% ===================================================
% ABSTRACT
% ===================================================
\begin{abstract}
%Given a system of partial differential equations, $F(t,x,\dot{x},p,u)=0$, where $x$ represents state variables, $p$ is the parameters, and $u$ are control variables, the process model is simultaneously solved for both $x_i$ and a set of sensitivity functions, $dx_i/dp_j$, overall times $t$.These sensitivity functions measure the influence of the parameter change on the model's output. As an example, the supercritical extraction process is presented. The impact of mass flow rate, pressure, and inlet temperature on the model's output is discussed. The sensitivity analysis results prove that the considered variables can affect the extraction's yield and be used as control variables in optimization problems. Moreover, the local sensitivity analysis results are analyzed from a phenomenological point of view to enhance understanding of the process model.
%Given a system of partial differential equations, $F(t,x,\dot{x},\Theta,u)=0$, where $x$ represents state variables, $\Theta$ are the parameters, and $u$ are control variables, we describe the supercritical extraction process. The process model describes a partially filled extractor with a fixed bed, which work under constant operating conditions. We assume that the flow is uniform across any cross-section, although the area available for the fluid phase can change along the extractor. We apply the concept of quasi-one-dimensional flow to mimic the modelling of a two-dimensional case.
%In this work, a distributed-parameter model, based on \citet{Reverchon1996} is used to describe a fluid-solid extraction process of caraway oil from caraway seeds with $CO_2$ as a solvent. The model parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient \textit{and saturation concentration} are obtained from parameter estimation. The parameters are estimated based on four experiments performed at $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar and $50^\circ C$/$300$ bar. Given yield data, the model-based parameter estimation uses the maximum likelihood estimation method under assumption of normal error.
This study aimed to investigate the supercritical extraction process of the caraway oil from caraway seeds from a partially filled extractor with a fixed bed operated under multiple constant operating conditions. The dataset collected from the experiments consists of four experiments conducted at different operating conditions: $40^\circ C$/$200$ bar, $50^\circ C$/$200$ bar, $40^\circ C$/$300$ bar, and $50^\circ C$/$300$ bar. A distributed-parameter model is used to describe the fluid-solid extraction process with $CO_2$ as a solvent. The concept of quasi-one-dimensional flow is applied to reduce the number of spatial dimensions. The flow is assumed to be uniform across any cross-section, although the area available for the fluid phase can vary along the extractor. This model requires parameters such as partition factor, internal diffusion coefficient, axial diffusion coefficient, saturation concentration, and an initial state estimate. These parameters are estimated from the Peng-Robinson equation of state and by applying the maximum likelihood estimation method on yield data under the normal error assumption.

@@

The quasi-one-dimensional compressible Navier-Stokes equations in Cartesian coordinates are given by Equations \ref{EQ: CompressibleEuler_1} to \ref{EQ: CompressibleEuler_3}. The derivation of these Equations are presented in Appendix \ref{CH: Gouverning equations}.

@@

{\footnotesize
	\begin{equation} \label{EQ: Pressure_Velocity}
		\cfrac{d v}{d {\color{orange}\rho_f}} = - \cfrac{d v}{d{\color{blue}P}} \cfrac{d {\color{blue}P}}{d {\color{orange}\rho_f}} = -\cfrac{1}{\rho v} \cfrac{d{\color{blue}P}}{d{\color{orange}\rho_f}} = -\cfrac{v}{{\color{orange}\rho_f}}
	\end{equation}
}

### 2023-06-27 SFE-article/Sections/Experiments.tex
07.05
27.06

	In order to solve the optimization problem presented in equation \ref{EQ: Optimization_formulation_MLE}, it is necessary to have knowledge of the dataset $Y(t)$, which was obtained by extracting caraway oil from caraway seeds. To prepare the seeds for extraction, they were first pre-treated using a Retsch SM 300 cutting mill to reduce their particle size to 1mm and break their outer shell. The moisture content of the seeds was then determined using an Infrared Moisture Analyser, which revealed an average moisture content of 4.83\% in the solid particles after grinding. Next, the density of the solid material was measured using a pycnometer, which can be found in appendix \ref{CH: Solid_Density_Measurment}. Finally, the material's porosity was also calculated to be 0.5, as explained in appendix \ref{CH: Porosity}.

@@

	After loading the material into the chamber, the extractor was pre-heated to the desired temperature. The outlet line was then closed, and the extraction chamber was filled with $CO_2$. The $CO_2$ was pumped and compressed until the desired operating pressure was reached. When the operating temperature and pressure were achieved, the outlet line was opened, allowing the solvent to flow through the system. The solvent extracted the essential oils from the solid material, and the resulting $CO_2$ and oil mixture flowed from the extractor to the separator. The separator operated at $50^\circ C$ and $50$ bar, with gaseous $CO_2$ exiting from the top and oil remaining at the bottom. The extraction time for each batch was 150 minutes, measured from the opening of the extractor outlet port. Every 5 minutes, the oil was drained from the separator, and its weight was measured. The resulting time series $Y(t)$ corresponds to the output of measurement equation \ref{Model_measurment_1} and can be used for parameter estimation.

@@

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
		\begin{tabular}{ccccc}
			\hline
			& $40 ^\circ C / 200$ bar & $50 ^\circ C / 200$ bar & $40 ^\circ C / 300$ bar & $50 ^\circ C / 300$ bar \\ 
			Time min & Mass g                  & Mass g                  & Mass g                  & Mass g                  \\ \hline
			0        & 0.0                     & 0.0                     & 0.0                     & 0.0                     \\
			5        & 1.7                     & 1.8                     & 1.1                     & 1.2                     \\
			10       & 5.7                     & 4.7                     & 4.7                     & 4.5                     \\
			15       & 10.9                    & 9.5                     & 10.8                    & 11.3                    \\
			20       & 17.8                    & 15.1                    & 20.6                    & 20.6                    \\
			25       & 24.9                    & 21.3                    & 32.1                    & 30.6                    \\
			30       & 30.3                    & 26.4                    & 40.5                    & 37.8                    \\
			35       & 34.0                    & 29.8                    & 46.7                    & 43.3                    \\
			40       & 37.5                    & 33.1                    & 50.5                    & 47.5                    \\
			45       & 40.1                    & 35.6                    & 53.1                    & 50.8                    \\
			50       & 43.3                    & 38.7                    & 55.5                    & 53.7                    \\
			55       & 45.5                    & 41.5                    & 57.6                    & 56.5                    \\
			60       & 47.6                    & 43.5                    & 59.3                    & 58.1                    \\
			65       & 49.2                    & 45.9                    & 60.9                    & 59.6                    \\
			70       & 50.8                    & 48.3                    & 62.2                    & 61.1                    \\
			75       & 52.6                    & 50.8                    & 63.7                    & 62.7                    \\
			80       & 54.2                    & 52.9                    & 64.9                    & 64.4                    \\
			85       & 55.1                    & 55.3                    & 66.6                    & 65.8                    \\
			90       & 56.2                    & 57.2                    & 67.6                    & 66.7                    \\
			95       & 57.3                    & 58.3                    & 68.9                    & 67.4                    \\
			100      & 58.3                    & 59.4                    & 69.7                    & 68.2                    \\
			105      & 59.4                    & 60.4                    & 70.6                    & 69.1                    \\
			110      & 60.2                    & 61.6                    & 71.2                    & 70.1                    \\
			115      & 61.4                    & 62.7                    & 71.9                    & 70.9                    \\
			120      & 62.5                    & 63.9                    & 72.3                    & 71.6                    \\
			125      & 63.3                    & 64.7                    & 72.8                    & 72.3                    \\
			130      & 64.0                    & 65.4                    & 73.2                    & 72.9                    \\
			135      & 64.8                    & 65.9                    & 73.6                    & 73.5                    \\
			140      & 65.5                    & 66.4                    & 74                      & 74                      \\
			145      & 66.2                    & 67.1                    & 74.3                    & 74.5                    \\
			150      & 66.8                    & 67.7                    & 74.6                    & 74.9              		 \\ \hline     
		\end{tabular} }
	\caption{Yield data}
	\label{tab: Yield_data}
	\end{table}

>>>>

	In order to solve the optimization problem presented in Equation \ref{EQ: Optimization_formulation_MLE}, it is necessary to have knowledge of the dataset ${\color{blue}Y}(t)$, which was obtained by extracting caraway oil from caraway seeds. To prepare the seeds for extraction, they were first pre-treated using a Retsch SM 300 cutting mill to reduce their particle size to 1mm and break their outer shell. The moisture content of the seeds was then determined using an Infrared Moisture Analyser, which revealed an average moisture content of 4.83\% in the solid particles after grinding. Next, the density of the solid material was measured using a pycnometer, which can be found in appendix \ref{CH: Solid_Density_Measurment}. Finally, the material's porosity was also calculated to be 0.5, as explained in appendix \ref{CH: Porosity}.

@@

	After loading the material into the extraction chamber, the extractor was pre-heated to the desired temperature. The outlet line was then closed, and the extraction chamber was filled with $CO_2$. The $CO_2$ was pumped and compressed until the desired operating pressure was reached. When the operating temperature and pressure were achieved, the outlet line was opened, allowing the solvent to flow through the system. The solvent extracted the essential oils from the solid material, and the resulting $CO_2$ and oil mixture flowed from the extractor to the separator. The separator operated at $50^\circ C$ and $50$ bar. The gaseous $CO_2$ exited from the top and was recycled to the $CO_2$ storage tank. The oil stayed at the bottom of the separator in the liquid form. The extraction time for each batch was 150 minutes, measured from the opening of the extractor outlet port. Every 5 minutes, the oil was drained from the separator, and its weight was measured. The resulting time series ${\color{blue}Y}(t)$ corresponds to the output of measurement Equation \ref{Model_measurment_1} and can be used for parameter estimation.

@@

%	\begin{table}[!h]
%		\centering
%		\adjustbox{max width=\columnwidth}{%
%		\begin{tabular}{ccccc}
%			\hline
%			& $40 ^\circ C / 200$ bar & $50 ^\circ C / 200$ bar & $40 ^\circ C / 300$ bar & $50 ^\circ C / 300$ bar \\ 
%			Time min & Mass g                  & Mass g                  & Mass g                  & Mass g                  \\ \hline
%			0        & 0.0                     & 0.0                     & 0.0                     & 0.0                     \\
%			5        & 1.7                     & 1.8                     & 1.1                     & 1.2                     \\
%			10       & 5.7                     & 4.7                     & 4.7                     & 4.5                     \\
%			15       & 10.9                    & 9.5                     & 10.8                    & 11.3                    \\
%			20       & 17.8                    & 15.1                    & 20.6                    & 20.6                    \\
%			25       & 24.9                    & 21.3                    & 32.1                    & 30.6                    \\
%			30       & 30.3                    & 26.4                    & 40.5                    & 37.8                    \\
%			35       & 34.0                    & 29.8                    & 46.7                    & 43.3                    \\
%			40       & 37.5                    & 33.1                    & 50.5                    & 47.5                    \\
%			45       & 40.1                    & 35.6                    & 53.1                    & 50.8                    \\
%			50       & 43.3                    & 38.7                    & 55.5                    & 53.7                    \\
%			55       & 45.5                    & 41.5                    & 57.6                    & 56.5                    \\
%			60       & 47.6                    & 43.5                    & 59.3                    & 58.1                    \\
%			65       & 49.2                    & 45.9                    & 60.9                    & 59.6                    \\
%			70       & 50.8                    & 48.3                    & 62.2                    & 61.1                    \\
%			75       & 52.6                    & 50.8                    & 63.7                    & 62.7                    \\
%			80       & 54.2                    & 52.9                    & 64.9                    & 64.4                    \\
%			85       & 55.1                    & 55.3                    & 66.6                    & 65.8                    \\
%			90       & 56.2                    & 57.2                    & 67.6                    & 66.7                    \\
%			95       & 57.3                    & 58.3                    & 68.9                    & 67.4                    \\
%			100      & 58.3                    & 59.4                    & 69.7                    & 68.2                    \\
%			105      & 59.4                    & 60.4                    & 70.6                    & 69.1                    \\
%			110      & 60.2                    & 61.6                    & 71.2                    & 70.1                    \\
%			115      & 61.4                    & 62.7                    & 71.9                    & 70.9                    \\
%			120      & 62.5                    & 63.9                    & 72.3                    & 71.6                    \\
%			125      & 63.3                    & 64.7                    & 72.8                    & 72.3                    \\
%			130      & 64.0                    & 65.4                    & 73.2                    & 72.9                    \\
%			135      & 64.8                    & 65.9                    & 73.6                    & 73.5                    \\
%			140      & 65.5                    & 66.4                    & 74                      & 74                      \\
%			145      & 66.2                    & 67.1                    & 74.3                    & 74.5                    \\
%			150      & 66.8                    & 67.7                    & 74.6                    & 74.9              		 \\ \hline     
%		\end{tabular} }
%	\caption{Yield data}
%	\label{tab: Yield_data}
%	\end{table}

	\begin{table}[!h]
		\centering
		\adjustbox{max width=\columnwidth}{%
			\csvautobooktabular{Figures/Results_estimation/Dataset.csv} }
		\caption{Yield data; cumulative amount of solute [g] collected every 5 min}
		\label{tab: Yield_data}
	\end{table}

### 2023-06-27 SFE-article/Sections/Model.tex
27.06

	\subsubsection{Continuity equation} \label{CH: Continuity}
	 %When the cross-sectional area of the channel $A_f$ is specified as a function of the void fraction ${\color{magenta}\phi}(z)$ (where ${\color{magenta}\phi}$ is the void fraction of the bed and $A$ is the cross-section of the empty extractor), the continuity equation takes the form: %\todo[]{assign a color of v and u}
	The discussed above quasi-one-dimensional assumption was used to derive the continuity equation for the fluid phase as presented in Appendix \ref{CH: Gouverning equations}. The cross-section change is taken into account by void fraction ${\color{magenta}\phi}(z)$, which depends on the spatial dimension. The equation \ref{EQ: Continuity_differential} shows the differential form, of the continuity equation:

@@

			The heat equation was introduced in the previous chapter through Equation \ref{EQ: CompressibleEuler_3} and in the appendix \ref{CH: Gouverning equations}

			{\footnotesize
			\begin{align}
				&\cfrac{\partial \left( {\color{orange}\rho_f}({\color{blue}T}(t,z),{\color{blue}P}(t)) {\color{blue}e}(t,z) A_f \right) }{\partial t} + \cfrac{\partial \left( {\color{orange}\rho_f}({\color{blue}T}(t,z),{\color{blue}P}(t)) A_f v {\color{blue}e}(t,z) \right)}{\partial {\color{blue}z}} \nonumber \\
				&= -{\color{blue}P}(t)\cfrac{\left( A_f v \right)}{\partial {\color{blue}z}} + \cfrac{\partial}{\partial {\color{blue}z}} \left( \cfrac{\partial {\color{blue}T}(t,z)}{\partial {\color{blue}z}} \right) 
			\end{align}
			}

@@

			If a gas is considered to be perfectly caloric (${\color{blue}e}(t,z)=C_v {\color{blue}T}(t,z)$), then the energy equation can be written in form of temperature. The perfectly caloric gas can be seen as the special case of a real gas, where the second term of the Equation \ref{EQ: internal_energy_def} goes to zero.

			For real gases, it is complicated to write the heat balance in terms of temperature, but it be can be used directly in the form of internal energy, as it is given by Equation \ref{EQ: CompressibleEuler_3}. In such a case, the temperature (which is used as the input to some other functions) needs to be recovered from the internal energy. A relation for the internal energy can be obtained from an equation of state. For Peng-Robinson, such a relation is given by Equation \ref{EQ: internal_energy_PR} as presented by \citet{Elliott2011}.

@@

			To solve Equation \ref{EQ: internal_energy_PR}, values of temperature, pressure, and density need to be known. If an equation of state is introduced, then only two out of three variables need to be obtained as the third one can be calculated, this can be represented as follow

@@

			If the value of internal energy ${\color{blue}e}(t,z)$ is known from the time evolution of the energy Equation \ref{EQ: CompressibleEuler_3}, and pressure is known from measurement, then the temperature can be reconstructed. A roootfinder can be used to find a value of temperature, which minimizes the difference between the value of internal energy coming from the time evolution and the output from Equation \ref{EQ: internal_energy_PR}. Such a procedure allow to find local temperature along spatial direction ${\color{blue}z}$ and needs to be repeated every time-step.

@@

			{\scriptsize
				\begin{equation}
					\begin{aligned}
					&{\color{blue}h}(t,z)-{\color{blue}h}(t,z)^{id}={\color{magenta}R}{\color{blue}T}(t,z) \\
					& \left[{\color{magenta}T_r}(Z-1)-2.078(1+{\color{magenta}\kappa} ){\sqrt {\alpha }}\ln \left(\cfrac{{\color{orange}Z}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)+\left( 1+\sqrt{2} \right){\color{orange}B}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)}{{\color{orange}Z}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)+\left( 1-\sqrt{2} \right){\color{orange}B}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)}\right)\right]
					\label{EQ:Enthalpy_PR}
				\end{aligned}
			\end{equation}				
			}

@@

		The pressure term in the energy equation, given by Equation \ref{EQ:Enthalpy_equation}, describes the change of the thermodynamic pressure with respect to time. As explained in Chapters \ref{CH: Gouverning equations} and \ref{CH:Low_Mach_chapter}, at low-Mach number conditions, the thermodynamic pressure is nearly constant in space due to the small pressure wave propagation that occurs at the speed of sound. Under such conditions, the term $\partial P/\partial t$ can be approximated by an ordinary differential equation, which describes the instantaneous change of pressure in the system. The pressure in the system is considered a state variable, while the pressure in the new time-step is considered a control variable.

>>>>

	\subsubsection{Continuity equation} \label{CH: Continuity}
	 %When the cross-sectional area of the channel $A_f$ is specified as a function of the void fraction ${\color{magenta}\phi}(z)$ (where ${\color{magenta}\phi}$ is the void fraction of the bed and $A$ is the cross-section of the empty extractor), the continuity equation takes the form: %\todo[]{assign a color of v and u}
	The obtained above quasi-one-dimensional continuity equation (Equation \ref{EQ: CompressibleEuler_1}) is further modified by specifying a function ${\color{magenta}A_f}(z) = {\color{magenta}A_f\phi}(z)$ to take into account the change of the cross-section available for the fluid. The cross-section change is taken into account by void fraction ${\color{magenta}\phi}(z)$, which depends on the spatial dimension. The equation \ref{EQ: Continuity_differential} shows the differential form, of the continuity equation:

@@

			The heat equation was introduced in the previous chapter through Equation \ref{EQ: CompressibleEuler_3}, in the appendix \ref{CH: Gouverning equations}  as well as through Equation \ref{EQ: Heat_internal}

			{\footnotesize
			\begin{align} \label{EQ: Heat_internal}
				&\cfrac{\partial \left( {\color{orange}\rho_f}({\color{blue}T}(t,z),{\color{blue}P}(t)) {\color{blue}e}(t,z) A_f \right) }{\partial t} + \cfrac{\partial \left( {\color{orange}\rho_f}({\color{blue}T}(t,z),{\color{blue}P}(t)) A_f v {\color{blue}e}(t,z) \right)}{\partial {\color{blue}z}} \nonumber \\
				&= -{\color{blue}P}(t)\cfrac{\left( A_f v \right)}{\partial {\color{blue}z}} + \cfrac{\partial}{\partial {\color{blue}z}} \left( \cfrac{\partial {\color{blue}T}(t,z)}{\partial {\color{blue}z}} \right) 
			\end{align}
			}

@@

			Suppose a gas is considered to be perfectly caloric (${\color{blue}e}(t,z)={\color{magenta}C_v} {\color{blue}T}(t,z)$), then the energy equation can be written explicitly in the form of temperature. The perfectly caloric gas can be seen as the special case of a real gas, where the second term of Equation \ref{EQ: internal_energy_def} goes to zero and the heat capacity ${\color{magenta}C_v}$ is constant.

			For real gases, it is complicated to write the heat balance in terms of temperature, but it be can be used directly in the form of internal energy, as it is given by Equation \ref{EQ: CompressibleEuler_3}. In such a case, the temperature needs to be recovered from the internal energy. A relation for the internal energy can be obtained from an equation of state. For Peng-Robinson, such a relation is given by Equation \ref{EQ: internal_energy_PR} as presented by \citet{Elliott2011}.

@@

			To solve Equation \ref{EQ: internal_energy_PR}, temperature, pressure, and density values need to be known. If an equation of state is introduced, then only two out of three variables need to be obtained as the third one can be calculated; this can be represented as follow

@@

			If the value of internal energy ${\color{blue}e}(t,z)$ is known from the time evolution of the energy Equation \ref{EQ: CompressibleEuler_3}, and pressure is known from measurement, then the temperature can be reconstructed. A roootfinder can be used to find a value of temperature, which minimizes the difference between the value of internal energy coming from the time evolution (Equation \ref{EQ: Heat_internal}) and the output from Equation \ref{EQ: internal_energy_PR}. Such a procedure allow to find local temperature along spatial direction ${\color{blue}z}$ and needs to be repeated every time-step.

@@

			{\scriptsize
				\begin{equation}
					\begin{split}
					{\color{blue}h}(t,z)-{\color{blue}h}(t,z)^{id}={\color{magenta}R}{\color{blue}T}(t,z) \left[{\color{magenta}T_r}({\color{orange}Z}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)-1) \right. \\
					\left. -2.078(1+{\color{magenta}\kappa} ){\sqrt { {\color{orange}\alpha}\left({\color{blue}T}(t,z)\right) } } \ln \left(\cfrac{{\color{orange}Z}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)+\left( 1+\sqrt{2} \right){\color{orange}B}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)}{{\color{orange}Z}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)+\left( 1-\sqrt{2} \right){\color{orange}B}\left({\color{blue}T}(t,z), {\color{red}P}(t)\right)}\right)\right]
					\label{EQ:Enthalpy_PR}
				\end{split}
			\end{equation}				
			}

@@

		The pressure term in the energy equation, given by Equation \ref{EQ:Enthalpy_equation}, describes the change of the thermodynamic pressure with respect to time. As explained in Chapters \ref{CH: Gouverning equations} and \ref{CH:Low_Mach_chapter}, at low-Mach number conditions, the thermodynamic pressure is nearly constant in space due to the small pressure wave propagation that occurs at the speed of sound. Under such conditions, the term $\partial {\color{blue}P}/\partial t$ can be approximated by an ordinary differential equation, which describes the instantaneous change of pressure in the system. The pressure in the system is considered a state variable, while the pressure in the new time-step is considered a control variable.

### 2023-06-27 SFE-article/Sections/Parameter_estimation.tex
27.06

	The goal of parameter estimation is to obtain the "best" estimate of parameter set ${\color{magenta} \theta}$ (which is a subset of the parameter space ${\color{magenta}\Theta}$ containing all parameters of a model) based on the continuous observations ${\color{blue}Y}(t)$ or the discrete observations ${\color{blue}Y}(t_i)$. Conceptually, the unobservable error $\epsilon(t)$ is added to the deterministic model output, ${\color{blue}y}(t)$ (Equation \ref{Model_measurment_1}), to give the observable dependent variable ${\color{blue}Y}(t)$ (for example results of an experiments). For discrete observations, this can be expressed as:

@@

	Maximum likelihood estimation (MLE) is a statistical method used to estimate the parameters of a probability distribution based on observed data. The MLE works by finding the values of the parameters that maximize the likelihood function, which is the probability of observing the given data for a given set of parameter values. The MLE has desirable properties such as asymptotic efficiency and normality. Although the MLE has often been associated with the normal distribution for mathematical convenience, it can be applied to a wide range of probability distributions. The derivation of the likelihood function under assumption of the Gaussian distribution is presented in the Appendix \ref{CH: ML}. The final form of the objective function is presented by Equation \ref{EQ: Objective_function}:

@@

	Based on the first order optimality condition, the $\ln L$ can be maximized with respect to the vector ${\color{magenta} \theta}$ by equating to zero the partial derivatives of $\ln L$ with respect to each of the parameters:

	{\footnotesize
		\begin{align}\label{EQ: MLE_Basic_Solution}
			\cfrac{\partial \ln L}{\partial {\color{magenta} \theta}} &= \cfrac{\partial \sum_{i=1}^{n} \ln p\left( y(t_i) | {\color{magenta} \theta} \right) }{\partial {\color{magenta} \theta}} = 0 
	\end{align} }

	Solution of Equations \ref{EQ: MLE_Basic_Solution} yield the desired estimates $\hat{{\color{magenta} \theta}}$. For some models, these equations can be explicitly solved for $\hat{{\color{magenta} \theta}}$ but in general no closed-form solution to the maximization problem is known or available, and an maximum likelihood estimator can only be found via numerical optimization.

>>>>

	The goal of parameter estimation is to obtain the "best" estimate of parameter set ${\color{magenta} \theta}$ (which is a subset of the parameter space ${\color{magenta}\Theta}$ containing all parameters of a model) based on the continuous observations ${\color{blue}Y}(t)$ or the discrete observations ${\color{blue}Y}(t_i)$. Conceptually, the unobservable error $\epsilon(t)$ is added to the deterministic model output, ${\color{blue}y}(t)$ (Equation \ref{Model_measurment_1}), to give the observable dependent variable ${\color{blue}Y}(t)$ (for example results of an experiment). For discrete observations, this can be expressed as:

@@

	Maximum likelihood estimation (MLE) is a statistical method used to estimate the parameters of a probability distribution based on observed data. The MLE works by finding the values of the parameters that maximize the likelihood function, which is the probability of observing the given data for a given set of parameter values. The MLE has desirable properties such as asymptotic efficiency and normality. Although the MLE has often been associated with the normal distribution for mathematical convenience, it can be applied to a wide range of probability distributions. The derivation of the likelihood function under the assumption of the Gaussian distribution is presented in Appendix \ref{CH: ML}. The final form of the objective function is presented by Equation \ref{EQ: Objective_function}:

@@

	Based on the first order optimality condition, the $\ln L$ can be maximized with respect to the vector ${\color{magenta} \theta}$ and equating to zero the partial derivatives of $\ln L$:

	{\footnotesize
		\begin{align}\label{EQ: MLE_Basic_Solution}
			\cfrac{\partial \ln L}{\partial {\color{magenta} \theta}} &= \cfrac{\partial \sum_{i=1}^{n} \ln p\left( {\color{blue}y}(t_i) | {\color{magenta} \theta} \right) }{\partial {\color{magenta} \theta}} = 0 
	\end{align} }

	Solution of Equations \ref{EQ: MLE_Basic_Solution} yield the desired estimates $\hat{{\color{magenta} \theta}}$. For some models, these equations can be explicitly solved for $\hat{{\color{magenta} \theta}}$ but in general, no closed-form solution to the maximization problem is known or available, and a maximum likelihood estimator can only be found via numerical optimization.
